On resolution sensitivity in the Community Atmosphere Model

### 2019-09-14 resolution.tex
edits

\title{Parameterized convection, stratiform clouds and resolution sensitivity in the Community Atmosphere Model}

@@

Observations show that $h=\frac{1}{3}$ for horizontal scales on the order of $100$ km and less \citep[hereafter referred to as the {\em{mesoscale}};][]{L1999JFM,CL2001JGR}. This value for $h$ is also supported by the slope of the kinetic energy spectrum $-\beta$, which can be related to the second-order structure function through the Weiner–Khinchine theorem, $\beta = \left( 2h+1 \right)$, valid in the range $1<\beta<3$ \citep[satisfies stationarity;][]{DETAL1996JAS}. For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is true for the kinetic energy spectrum in observations \citep{NG1985JAS,CETAL1999JGR} and models \citep[e.g.,][]{THO2006GRL,SPKS2014JAS} at the mesoscale. \cite{RETAL2016CD} propose that the overwhelming support for $h=\frac{1}{3}$ at the mesoscale provides an emergent constraint for $n=-\frac{2}{3}$ in equation~\ref{eq:alpha}.

@@

Another robust response of the CAM/CCM lineage to resolution is an increase in stratiform precipitation rates at the expense of parameterized convective precipitation rates. The resolution dependent partitioning between the two different precipitation routines is shown in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies. The tendency for precipitation rates to shift from the convection scheme to the stratiform scheme with resolution has been documented in other models \citep{PS2002CD,RETAL2016CD,TETAL2018CD}, but none have provided a satisfactory explanation for this sensitivity. The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the practice of reducing $\Delta t_{phys}$ with resolution should by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convective precipitation rates with resolution.

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=20pc,angle=0]{figs/cam-history.pdf}\\
\end{center}
\caption{Bar-graph of the convective (solid) and stratiform (white) climatological precipitation rates in prior CAM/CCM convergence studies. Each window contains a single convergence study, with identical x-axis; the approximate grid resolution. Colors indicate the model configuration; January ensemble (black) and aqua-planet configurations with SST profiles $QOBS$ (blue) and $CNTL$ (red) in \cite{NH2000ASL}. Studies included in this figure are \cite{KW1991JGR} (CCM1), \cite{WETAL1995CD} (CCM2), \cite{W2008TELLUS} (CAM3), \cite{RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM} (CAM4), \cite{ZetAl2014JCb} (CAM5) and this study (CAM6). CCM2* refers to the modified parameter experiment of \cite{WETAL1995CD}, where parameters vary with resolution to reduce the dependence of cloud fraction on resolution.}
\label{fig:cam-history}
\end{figure}

In this study, a convergence experiment using CAM, version 6 (CAM6; \url{https://ncar.github.io/CAM/doc/build/html/users_guide/index.html}) is carried out and analyzed in detail. It is shown that the resolution sensitivity of vertical velocities are well described with $n=-1$ in equation~\eqref{eq:alpha}, provided $\Delta t_{phys}$ is defined in a way that avoids large truncation errors across resolutions. The increase in stratiform precipitation rates with resolution are attributed to an increase in resolved moisture fluxes at cloud base, primarily due to the increase in upward vertical velocities with resolution. The corresponding increase in downward vertical velocities with resolution are shown to stabilize the mean state, reducing the activity of the convection scheme and associated precipitation rates with increasing resolution. The feedback of the resolved vertical motion on the physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution. Section 2 describes CAM6 and the details of the convergence experiment. Section 3 contains a thorough analysis of the CAM6 simulations and Section 4 provides some discussion and conclusions.

 \begin{table*}
 \caption{Experimental design and global mean climatologies. $\Delta x$ refers to the average equatorial grid-spacing.}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   Variable & $ne20$ & $ne30$ & $ne40$ & $ne60$ & $ne80$ & $ne120$ \\ 
   \hline
   $\Delta x$ (km) & 166.8 & 111.2 & 83.4 & 55.6 & 41.7 & 27.8 \\
   $\nu$ ($m^4/s$) & $1.5 \times 10^{15}$ & $4.0 \times 10^{14}$ & $1.5 \times 10^{14}$ & $4.0 \times 10^{13}$  & $1.5 \times 10^{13}$ & $4.0 \times 10^{12}$\\
    $\Delta t_{phys}$ (s) & 2700 & 1800 & 1350 & 900 & 675 & 450 \\
   Total Cloud Fraction & 0.844 & 0.835 & 0.824 & 0.810 & 0.804 & 0.800 \\ 
   Total Precipitable Water (mm) & 23.31& 23.01 & 22.62 & 22.25 & 21.93 & 21.72 \\
   Convective Precipitation (mm/day) & 1.91 & 1.83 & 1.68 & 1.47 & 1.29 & 1.08 \\
   Stratiform Precipitation (mm/day) & 1.26 & 1.42 & 1.60 & 1.85 & 2.05 & 2.22 \\      
 \hline
 \end{tabular}
 \label{tbl:table1}
 \end{table*}

@@

This study uses the spectral-element dynamical core option of Community Atmosphere Model \citep[CAM-SE;][]{DetAl2012IJHPCA}, coupled with a mass conserving, semi-Lagrangian advection method for accelerated multi-tracer transport \citep[CSLAM;][]{LTOUNGK2017MWR}, and dry-mass vertical coordinate with comprehensive treatment of moisture and energy \citep{LetAl2018JAMES}. The dry dynamics are solved using the high-order, momentum, mass and energy conserving spectral element method \citep{TF2010JCP}, with the elements defined by a cubed-sphere grid. The notation for the horizontal grid resolution is an `$ne$' followed by the number of elements making up an edge of one cubed-sphere face, e.g., $ne30$. Hyper-viscous $\nabla^{4}$ explicit numerical dissipation is applied to temperature, dry pressure thickness, rotational and divergent winds \citep{LetAl2018JAMES}. CSLAM tracer transport uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.

@@

The convergence experiment is performed in an aqua-planet configuration \citep{NH2000ASL,MWO2016JAMES}, an all ocean planet with fixed, zonally symmetric sea surface temperatures modeled after present day Earth \citep[$QOBS$ in][]{NH2000ASL}. The aqua-planets are in a perpetual equinox, and aerosols are largely absent from the simulations. Each simulation is ran for one simulated year. Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}. In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing resolution sensitivity in this study. All analyses exclude the first month of the simulations, and are computed on their native grids unless otherwise stated.

@@

The PDF's may be scaled to the highest-resolution resolution grid through $ \alpha P (\omega / \alpha)$, where $P(x)$ is the PDF, and $\alpha$ the scale factor from equation~\ref{eq:alpha} with $\Delta x_0$ set to the $ne120$ grid-spacing. Figure~\ref{fig:2pdf}b shows the scaled PDF's for a power law exponent $n=-1$ in $\Delta x$. The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law exponent $n=-1$ explains to first-order the variation in vertical velocity with resolution as shown by the aqua-planet simulations.

Changes to the vertical velocity field can be further understood through decomposing the mass weighted vertical mean $ \langle \omega \rangle$ into upward and downward components,
\begin{equation}
\langle \omega \rangle =\langle f_{u} \rangle \, \langle \omega_{u} \rangle + \langle f_{d} \rangle \, \langle \omega_{d} \rangle, \label{eq:omega}
\end{equation}
where $\langle f_x \rangle$ and $\langle \omega_x \rangle$ refers to the vertical mass fraction $ \left( \frac{\int dp_x}{\int dp} \right)$ and the $x$ component of the mass weighted vertical mean of $\omega$ $ \left( \frac{\int \omega_x dp_x}{\int dp_x} \right)$, respectively, subscript $u$ refers to upward motion and $d$, downward motion.

@@

 \begin{table}
 \caption{Fractional contribution of latitude bands $\pm 10^{\circ}$ and $\pm 15^{\circ}$ to changes in global mean precipitation with resolution. The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc... All differences are computed after conservative remapping to a common $ne20$ grid.}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   Variable & $ne30$ & $ne40$ & $ne60$ & $ne80$ & $ne120$ \\ 
   \hline
   $\pm 10^{\circ}$ ($17.6\%$ of global area) \\
   Convective Precipitation & -0.58 & 0.62 & 0.66 & 0.72 & 0.70 \\
   Stratiform Precipitation & 0.55 & 0.63 & 0.69 & 0.67 & 0.41 \\ 
   \hline
   $\pm 15^{\circ}$ ($25.8\%$ of global area) \\
   Convective Precipitation & 0.22 & 0.75 & 0.73 & 0.79 & 0.72 \\
   Stratiform Precipitation & 0.46 & 0.64 & 0.71 & 0.70 & 0.49 \\      
 \hline
 \end{tabular}
 \label{tbl:table2}
 \end{table}

@@

$\overline{P_s}$ can be expressed as the double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
\begin{equation}
\overline{P_{s}} = \sum_i \sum_j f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right) \label{eq:pdecomp}
\end{equation}
where the subscript $850$ is dropped from $\omega$ and $q$ for brevity. $f_s$ is a measure of the occurrence of a particular combination $\left( \omega_i , q_j \right)$ in a simulation, and $M_s$ the mean stratiform precipitation rate associated with the combination $\left( \omega_i , q_j \right)$. Their product $f_s M_s$ is then the contribution of a particular combination of $\left( \omega_i , q_j \right)$ to $\overline{P_s}$.

@@

While the increase in magnitude of vertical velocities with resolution clearly drives an increase in moisture fluxes and stratiform precipitation rates with resolution, it is not clear why their is a reduction parameterized deep convective precipitation rates with resolution, in all versions of the model (Figure~\ref{fig:cam-history}). Curiously though, there is an excellent negative correlation (Pearson's R-value = 0.99, N = 27) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}). At any grid-point and time-step, $FREQZM$ is a binary variable: 1 if the ZM scheme is active, 0 if it is not. Time mean $FREQZM$ is therefore the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg. The regression indicates that model simulations with greater subsidence also have less convective activity.

@@

To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then through fixing only the temperature to the $ne20$ profile. Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field. Moisture changes with resolution do not contribute to the reduction in dilute CAPE with resolution, with the exception that the 20 J/kg reduction in dilute CAPE of ascending regions between the $ne80$ and $ne120$ runs (Figure~\ref{fig:cape}a) is entirely due to changes in moisture (not shown).

@@

Figure~\ref{fig:cape-subt}a shows the dilute CAPE values computed from mean temperature and moisture of subsiding and ascending regions in the $\pm \left( 10^{\circ}-30^{\circ} \right) $ latitude bands, hereafter referred to as the subtropics. The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), and their dilute CAPE  values vary much less with resolution ($\sim 20$ J/kg, compared with $\sim 80$ J/kg across all resolutions in the deep tropics). Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but with temperature changes playing a larger role (not shown). This analysis suggests that increasing subsidence with resolution (Figure~\ref{fig:4zonal}a) increases the stability of the mean state, since both ascending/descending regimes become more stable with resolution, and reducing climatological $FREQZM$ (Figure~\ref{fig:4zonal}b) despite the poor relationship between instantaneous subsidence and $FREQZM$ in the subtropics, as shown by the logistic regression (Figure~\ref{fig:4zonal}c).

@@

Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of physical parameterizations compatible with the highly flexible grid structures now being used to maximize efficiency on modern day supercomputers (scale-aware physics). This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}). As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution. Prior studies have established that the drying and reduction in cloudiness with resolution is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advects dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,HR2017JCLIM}. This study aims to then understand why and how vertical velocities change with resolution, and why convective precipitation decreases at the expense of stratiform precipitation with increasing resolution.

@@

The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind. \cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations, it is an emergent constraint for $n=-\frac{2}{3}$. This is an intriguing proposal and the authors don't seek to dismiss its potential value. But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km, since CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}. The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem. For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is significantly shallower than the slope observed near the effective resolution for all grid resolutions in this study (Figure~~\ref{fig:2ke}).

@@

An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, balanced by an increase in stratiform precipitation rates. Through decomposing the stratiform precipitation into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved mass fluxes arising from the increase in resolved vertical velocities with resolution. At higher resolutions, larger stratiform precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate, with smaller magnitude precipitation rates contributing less. The increasing occurrence of larger magnitude vertical velocities and stratiform precipitation rates at higher resolution is at odds with a similar analysis by \cite{TETAL2018CD} using a similar model. The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux) which complicated their calculation.

@@

Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE \citep[DCAPE;][]{XZ2000JGR,Z2002JGR}. Subsidence warming is represented in the DCAPE budget as negative the vertical advection of dry static energy term, since warming the environment reduces DCAPE, and simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F. Song, personal communication). The magnitude of negative, or downward $w$ increases everywhere in the model with increasing resolution, opposing the growth of dilute CAPE and reducing the activity of the ZM scheme with increasing resolution.

>>>>

\title{On resolution sensitivity in the Community Atmosphere Model}

@@

Observations show that $h=\frac{1}{3}$ for horizontal scales on the order of $100$ km and less \citep[hereafter referred to as the {\em{mesoscale}};][]{L1999JFM,CL2001JGR}. This value for $h$ is also supported by the slope of the kinetic energy spectrum $-\beta$, which can be related to the second-order structure function through the Weiner–Khinchine theorem, $\beta = \left( 2h+1 \right)$, valid in the range $1<\beta<3$ \citep[satisfies stationarity;][]{DETAL1996JAS}. For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is true for the kinetic energy spectrum in observations \citep{NG1985JAS,CETAL1999JGR} and models at the mesoscale \citep[e.g.,][]{THO2006GRL,SPKS2014JAS}. \cite{RETAL2016CD} propose that the overwhelming support for $h=\frac{1}{3}$ at the mesoscale provides an emergent constraint for $n=-\frac{2}{3}$ in equation~\ref{eq:alpha}.

@@

Another robust response of the CAM/CCM lineage to resolution is an increase in stratiform precipitation rates at the expense of parameterized convective precipitation rates. Stratiform precipitation is sometimes referred to as grid-scale precipitation because unlike parameterized convection, it condenses moisture locally without any transport by unresolved eddies. The resolution dependent partitioning between the two different precipitation routines is shown in Figure~\ref{fig:cam-history}, which is a bar-graph of the climatological, global mean stratiform and convective precipitation rates in prior CAM/CCM convergence studies. The tendency for precipitation rates to shift from the convection scheme to the stratiform scheme with resolution has been documented in other models \citep{PS2002CD,RETAL2016CD,TETAL2018CD}, but none have provided a convincing explanation for this sensitivity. The studies of \cite{KW1991JGR}, \cite{WETAL1995CD} and \cite{W2013QJRMS} indicate that the practice of reducing $\Delta t_{phys}$ with resolution should by itself reduce the convective precipitation rates, however Figure~\ref{fig:cam-history} (top row) indicates that convergence studies with fixed $\Delta t_{phys}$ still show a reduction in convective precipitation rates with resolution.

\begin{figure}[t]
\begin{center}
\noindent\includegraphics[width=20pc,angle=0]{figs/cam-history.pdf}\\
\end{center}
\caption{Bar-graph of the convective (solid) and stratiform (white) climatological precipitation rates in prior CAM/CCM convergence studies. Each window contains a single convergence experiment, with identical x-axis; the approximate grid resolution. Colors indicate the model configuration; January ensemble (black) and aqua-planet configurations with SST profiles $QOBS$ (blue) and $CNTL$ (red) in \cite{NH2000ASL}. Studies included in this figure are \cite{KW1991JGR} (CCM1), \cite{WETAL1995CD} (CCM2), \cite{W2008TELLUS} (CAM3), \cite{RETAL2013JCLIM,ZetAl2014JCb,HR2017JCLIM} (CAM4), \cite{ZetAl2014JCb} (CAM5) and this study (CAM6). CCM2* refers to the modified parameter experiment of \cite{WETAL1995CD}, where parameters vary with resolution to reduce the dependence of cloud fraction on resolution.}
\label{fig:cam-history}
\end{figure}

In this study, a convergence experiment using CAM, version 6 (CAM6; \url{https://ncar.github.io/CAM/doc/build/html/users_guide/index.html}) is carried out and analyzed in detail. It is shown that the resolution sensitivity of vertical velocities are well described with $n=-1$ in equation~\eqref{eq:alpha}, provided $\Delta t_{phys}$ is defined in a way that avoids large truncation errors across resolutions. The increase in stratiform precipitation rates with resolution are attributed to an increase in resolved moisture fluxes at cloud base, primarily due to the increase in upward vertical velocities with resolution. The corresponding increase in downward vertical velocities with resolution are shown to stabilize the mean state, reducing the activity of the convection scheme and associated precipitation rates with increasing resolution. The feedback of the resolved vertical motion on the model physics indicates that the root cause of resolution sensitivity in CAM arises from the sensitivity of resolved dynamical modes to native grid resolution. Section 2 describes CAM6 and the details of the convergence experiment. Section 3 contains a thorough analysis of the CAM6 simulations and Section 4 provides some discussion and conclusions.

 \begin{table*}
 \caption{Experimental design and global mean climatologies. $\Delta x$ refers to the average equatorial grid-spacing.}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   Variable & $ne20$ & $ne30$ & $ne40$ & $ne60$ & $ne80$ & $ne120$ \\ 
   \hline
   $\Delta x$ (km) & 166.8 & 111.2 & 83.4 & 55.6 & 41.7 & 27.8 \\
   $\nu$ ($m^4/s$) & $1.5 \times 10^{15}$ & $4.0 \times 10^{14}$ & $1.5 \times 10^{14}$ & $4.0 \times 10^{13}$  & $1.5 \times 10^{13}$ & $4.0 \times 10^{12}$\\
    $\Delta t_{phys}$ (s) & 2700 & 1800 & 1350 & 900 & 675 & 450 \\
   Total Cloud Fraction & 0.844 & 0.835 & 0.824 & 0.810 & 0.804 & 0.800 \\ 
   Total Precipitable Water (mm) & 23.31& 23.01 & 22.62 & 22.25 & 21.93 & 21.72 \\
   Convective Precipitation, $P_c$ (mm/day) & 1.91 & 1.83 & 1.68 & 1.47 & 1.29 & 1.08 \\
   Stratiform Precipitation, $P_s$ (mm/day) & 1.26 & 1.42 & 1.60 & 1.85 & 2.05 & 2.22 \\      
 \hline
 \end{tabular}
 \label{tbl:table1}
 \end{table*}

@@

This study uses the spectral-element dynamical core option of Community Atmosphere Model \citep[CAM-SE;][]{DetAl2012IJHPCA}, coupled with a mass conserving, semi-Lagrangian advection method for accelerated multi-tracer transport \citep[CSLAM;][]{LTOUNGK2017MWR}, and dry-mass vertical coordinate with comprehensive treatment of moisture and energy \citep{LetAl2018JAMES}. The dry dynamics are solved using the high-order, momentum, mass and energy conserving spectral element method \citep{TF2010JCP}, with the elements defined by a cubed-sphere grid. The notation for horizontal grid resolution is an `$ne$' followed by the number of elements making up an edge of one cubed-sphere face, e.g., $ne30$. Hyper-viscous $\nabla^{4}$ damping is applied to temperature, dry pressure thickness, rotational and divergent winds \citep{LetAl2018JAMES}. CSLAM tracer transport uses a finite volume grid constructed from the cubed-sphere of elements, and contains the same degrees of freedom as the dry dynamics.

@@

The convergence experiment is performed in an aqua-planet configuration \citep{NH2000ASL,MWO2016JAMES}, an all ocean planet with fixed, zonally symmetric sea surface temperatures modeled after present day Earth \citep[$QOBS$ in][]{NH2000ASL}. The aqua-planets are in a perpetual equinox, and aerosols are largely absent from the simulations. Each simulation is ran for one simulated year. Six different horizontal grids are used in this study, which are provided in Table~\ref{tbl:table1}. In addition to the six simulations used in the convergence experiment, an ensemble of 24 simulations containing different model parameters (e.g., using the lower resolution physics grid) and across different resolutions are ran for one year in order to increase confidence in assessing the resolution sensitivity in this study. All analyses exclude the first month of the simulations, which is sufficient to avoid the spin-up period; the model comes into global energy balance after about ten days (not shown). All calculations are performed on their native grids unless otherwise stated.

@@

The PDF's may be scaled to the highest-resolution, $ne120$ grid through $ \alpha P (\omega / \alpha)$, where $P(x)$ is the PDF and $\alpha$ the scale factor from equation~\ref{eq:alpha} with $\Delta x_0$ set to the $ne120$ grid-spacing. Figure~\ref{fig:2pdf}b shows the scaled PDF's for a power law exponent $n=-1$ in $\Delta x$. The scaled PDF's all collapse onto the high-resolution reference, indicating that the power-law exponent $n=-1$ explains to first-order the variation in vertical velocity with resolution as shown by the aqua-planet simulations.

Changes to the vertical velocity field can be further understood through decomposing the mass weighted vertical mean $ \langle \omega \rangle$ into upward and downward components,
\begin{equation}
\langle \omega \rangle =\langle f_{u} \rangle \, \langle \omega_{u} \rangle + \langle f_{d} \rangle \, \langle \omega_{d} \rangle, \label{eq:omega}
\end{equation}
where $\langle f_x \rangle$ and $\langle \omega_x \rangle$ refers to the vertical mass fraction $ \left( \frac{\int dp_x}{\int dp} \textrm{, with pressure } p \right)$ and the $x$ component of the mass weighted vertical mean of $\omega$ $ \left( \frac{\int \omega_x dp_x}{\int dp_x} \right)$, respectively, subscript $u$ refers to upward motion and $d$, downward motion.

@@

 \begin{table}
 \caption{Fractional contribution of latitude bands $\pm 10^{\circ}$ and $\pm 15^{\circ}$ to changes in global mean stratiform and parameterized convective precipitation with resolution. The grid headers refer to differences with respect to the next lowest grid resolution, e.g., $ne30 = ne30-ne20$, $ne40=ne40-ne30$, etc... All differences are computed after conservative remapping to a common $ne20$ grid.}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   Variable & $ne30$ & $ne40$ & $ne60$ & $ne80$ & $ne120$ \\ 
   \hline
   $\pm 10^{\circ}$ ($17.6\%$ of global area) \\
   Convective Precipitation, $P_c$ & -0.58 & 0.62 & 0.66 & 0.72 & 0.70 \\
   Stratiform Precipitation, $P_s$ & 0.55 & 0.63 & 0.69 & 0.67 & 0.41 \\ 
   \hline
   $\pm 15^{\circ}$ ($25.8\%$ of global area) \\
   Convective Precipitation, $P_c$ & 0.22 & 0.75 & 0.73 & 0.79 & 0.72 \\
   Stratiform Precipitation, $P_s$ & 0.46 & 0.64 & 0.71 & 0.70 & 0.49 \\      
 \hline
 \end{tabular}
 \label{tbl:table2}
 \end{table}

@@

$\overline{P_s}$ can be expressed as the double sum of the product of the time mean magnitude $M_s$ and time mean spatial frequency $f_s$, over $\omega_{850}$ and $q_{850}$ space,
\begin{equation}
\overline{P_{s}} = \sum_i \sum_j f_s \left( \omega_i , q_j \right) M_s \left( \omega_i , q_j \right) \label{eq:pdecomp}
\end{equation}
where the subscript $850$ is dropped from $\omega$ and $q$ for brevity. $f_s$ is a measure of the occurrence of a particular combination $\left( \omega_i , q_j \right)$ in a simulation, and $M_s$ the mean stratiform precipitation rate associated with the combination $\left( \omega_i , q_j \right)$. Their product $f_s M_s$ is then the contribution of the combination $\left( \omega_i , q_j \right)$ to $\overline{P_s}$.

@@

While the increase in magnitude of vertical velocities with resolution clearly drives an increase in moisture fluxes and stratiform precipitation rates with resolution, it is not clear why there is a reduction in parameterized deep convective precipitation rates with resolution, in all versions of the model (Figure~\ref{fig:cam-history}). Curiously though, there is an excellent negative correlation (Pearson's R-value = 0.99, N = 27) between the global mean, climatological $\langle f_{d} \rangle \, \langle \omega_{d} \rangle$ and a measure of the activity of the \cite{ZM1995AO} deep convection scheme (referred to as the {\em{ZM scheme}} hereafter), global mean, climatological $FREQZM$ (Figure~\ref{fig:corr}). At any grid-point and time-step, $FREQZM$ is a binary variable: 1 if the ZM scheme is active, 0 if it is not. Time mean $FREQZM$ is therefore the fraction of the model time that the ZM scheme is triggered, i.e., dilute CAPE exceeds $\geq 70$ J/kg. The regression indicates that model simulations with greater subsidence also have less convective activity.

@@

To isolate the relative importance of temperature or moisture on the systematic reduction in dilute CAPE with resolution for both ascending/descending regimes, dilute CAPE is recomputed for all resolutions, but through fixing the moisture profiles to the lowest resolution $ne20$ profile, and then through fixing only the temperature to the $ne20$ profile. Figure~\ref{fig:cape}d shows the influence of changing temperature profile with resolution, with moisture fixed to the $ne20$ profile, on dilute CAPE of ascending/descending grid columns, and illustrates that the systematic reduction in dilute CAPE with resolution in both regimes is primarily from changes to the temperature field. Moisture changes with resolution do not contribute to the reduction in dilute CAPE with resolution, with the exception that the 20 J/kg reduction in dilute CAPE of ascending regions between the $ne80$ and $ne120$ runs (Figure~\ref{fig:cape}a) is entirely due to changes in moisture (not shown).

@@

Figure~\ref{fig:cape-subt}a shows the dilute CAPE values computed from mean temperature and moisture of subsiding and ascending regions in the $\pm \left( 10^{\circ}-30^{\circ} \right) $ latitude bands, hereafter referred to as the subtropics. The spread in dilute CAPE between ascending and descending grid columns is much smaller than for the deep tropics (Figure~\ref{fig:cape}a), and their dilute CAPE  values vary much less with resolution ($\sim 20$ J/kg, compared with $\sim 80$ J/kg across all resolutions in the deep tropics). Through recomputing dilute CAPE and fixing the temperature or moisture profiles to $ne20$ values, the reduction in dilute CAPE with resolution is attributed to both temperature and moisture, but as in the deep tropics, temperature changes have a larger influence (not shown). This analysis suggests that increasing subsidence with resolution (Figure~\ref{fig:4zonal}a) increases the stability of the mean state, since both ascending/descending regimes become more stable with resolution, and reducing climatological $FREQZM$ (Figure~\ref{fig:4zonal}b) despite the poor relationship between instantaneous subsidence and $FREQZM$ in the subtropics, as shown by the logistic regression (Figure~\ref{fig:4zonal}c).

@@

Establishing a complete understanding of resolution sensitivity in atmospheric general circulation models (AGCMs) is crucial, since inevitable responses to native grid resolution cannot be ignored in the pursuit of physical parameterizations compatible with the highly flexible grid structures now being used to maximize efficiency on modern day supercomputers (i.e., scale-aware physics). This study analyzes a convergence experiment in an aqua-planet configuration using the Community Atmosphere Model (CAM), with the spectral-element dynamical core option coupled to the accelerated multi-tracer transport scheme (CAM-SE-CSLAM), and version 6 physics (CAM6), in order to understand longstanding convergence issues that have persisted within the CAM lineage since its inception (see Figure~\ref{fig:cam-history}). As with most versions of CAM, in CAM6 the atmosphere becomes drier and less cloudy with increasing resolution, and parameterized convective precipitation decreases while stratiform precipitation increases with resolution. Prior studies have established that the drying and reduction in cloudiness with resolution is a result of increases in the magnitude of vertical velocities with resolution, which more efficiently advects dry air aloft, downward \citep{KW1991JGR,WETAL1995CD,HR2017JCLIM}. This study aims to then understand why and how vertical velocities change with resolution, and why convective precipitation decreases at the expense of stratiform precipitation with increasing resolution.

@@

The $n=-1$ power law scaling is potentially in conflict with the scaling proposed by \cite{RETAL2016CD}, $n=h-1$, with $h$ being half the exponent of the second-order structure function $\Delta x^{2h}$ of the horizontal wind. \cite{RETAL2016CD} propose that since the $h=\frac{1}{3}$ is supported by observations, it is an emergent constraint for $n=-\frac{2}{3}$. This is an intriguing proposal and the authors don't seek to dismiss its potential value. But it is suggested that the overwhelming support for the $h=\frac{1}{3}$ exponent in the second-order structure function is only relevant for horizontal scales on the order of $100$ km and less \citep{L1999JFM,CL2001JGR}, which is only marginally resolved in the highest resolution simulation $\Delta x=27.8$ km ($ne120$), and CAM-SE-CSLAM has an effective resolution in the range of $5-10\Delta x$ \citep{HETAL2019JAMES}. The nonexistence of $h=\frac{1}{3}$ in the $\Delta x=27.8$ km simulation is verified through analyzing the slope of kinetic energy spectrum $-\beta$, which is in duality with the second-order structure function through the Weiner–Khinchine theorem. For $h=\frac{1}{3}$, $\beta=\frac{5}{3}$, which is significantly shallower than the slope observed near the effective resolution for all grid resolutions in this study (Figure~~\ref{fig:2ke}).

@@

An increase in resolved moisture fluxes through cloud base occurs with increasing resolution, and balanced by an increase in stratiform precipitation rates. Through decomposing the stratiform precipitation into the components of the moisture flux, it is shown that the increase in stratiform precipitation rates with resolution is primarily due to the increase in resolved mass fluxes arising from the increase in resolved vertical velocities with resolution. At higher resolutions, larger stratiform precipitation rates increasingly contribute to the increase in climatological stratiform precipitation rate, with smaller magnitude precipitation rates contributing less. The increasing occurrence of larger magnitude vertical velocities and stratiform precipitation rates at higher resolution is at odds with a similar analysis by \cite{TETAL2018CD} using a similar model. The authors suggests that their use of daily averaged fields obscured the covariance between vertical velocity and moisture (i.e., moisture flux) and complicated their calculation.

@@

Subsidence warming falls out of the component of the dilute CAPE budget due to the advection of dry static energy and moisture by the environment, commonly referred to as dynamic CAPE \citep[DCAPE;][]{XZ2000JGR,Z2002JGR}. Subsidence warming is represented in the DCAPE budget as negative the vertical advection of dry static energy term, since warming the environment reduces DCAPE, and simplifies to $\frac{w}{c_{pd}} \frac{\partial gz}{\partial z} = \frac{g}{c_{pd}} w$, directly proportional to the vertical velocity $w$ by the factor $\frac{g}{c_{pd}}$, with $g$ the acceleration of gravity and $c_{pd}$ the specific heat capacity of dry air (F. Song, personal communication). The magnitude of negative, or downward, $w$ increases everywhere in the model with increasing resolution, opposing the growth of dilute CAPE and reducing the activity of the ZM scheme with increasing resolution.
